### 2014-05-24 patchy-selection-paper.tex
added applications

  new mutation: \eqref{eqn:mutrate};

@@

Taking these approximations at face value, 
and assuming both rates are small, 
the time until the first of the two patches is colonized by the $B$
allele will be approximately exponentially distributed with mean $1/(2
\mutrate)$.
Following this, the time until the second patch is subsequently colonized 
(via either migration or new mutation) 
will be exponentially distributed with mean $1/(\mutrate+\migrate)$.
Both these rates scale linearly with population density ($\rho$) 
and the probability of establishment of a rare allele ($p_e\approx 2 s_b/\xi^2$),
so that the overall time to adaptation is robustly predicted to increase with offspring variance ($\xi^2$)
and decrease with population density and benefical selection coefficient.
Furthermore, the probability that the second adaptation is a new mutation,
i.e.\ the probability of parallel adapatation, is in $d=1$ ($d=2$ only differs slightly)
\begin{equation}
  \frac{\mutrate}{\mutrate+\migrate(r)} = \frac{ A \mu \alpha }{2 A \mu \alpha + \exp\left(- \sqrt{2 s_m} r / \sigma \right) },
\end{equation}
so that the probability of parallel mutation should increase
approximately logistically with the distance between the patches, at rate $\sqrt{s_m} /\sigma$.

@@

If not yet adapted, a patch of area $A$ will acquire the adaptation by new mutation at rate $2 s_p \mu \rho A/\xi^2$ (equation \eqref{eqn:mutrate}).
Suppose that patches $1, \ldots, k$ are already adapted,
and that these patches are at distances $r_1, \ldots, r_k$ away from this unadapted patch.
The total rate of adaptation through migration, from equation \eqref{eqn:migrate}, is
\begin{equation}
  \migrate(r_i) = C \frac{ \rho s_p }{\xi^2 s_m} \sum_{i=1}^{k} \left(r_i \sqrt{2 s_m} /\sigma \right)^{-(d-1)/2} \exp\left(- r_i \sqrt{2 s_m} /\sigma\right),
\end{equation}
and if the patch adapts through migration, the probability the adapted allele 
comes from patch $i$ is 
\begin{align}
  \frac{\left(r_i \right)^{-(d-1)/2} \exp\left(- r_i \sqrt{2 s_m}
  /\sigma\right)} {\sum_{j=1}^{k}  \left(r_i \right)^{-(d-1)/2} \exp\left(- r_i \sqrt{2 s_m}
    /\sigma\right) } .
\end{align}
These specify the transition rates of the process.

@@

\gc{add some
  sentences of intuition of how parameters play into haplotype length
  here. is it really $\sigma \sqrt{s_m}$ in \eqref{eqn:haplotype_length} and not $\sqrt{s_m}/sigma$???}

\gc{Perhaps we should separate the above bit into things that are
  useful about $\tau$ and put those up by  \eqref{eqn:borodinresult} and
  things that are useful about $L$ here? We should also turn some of
  this into numbered equations. I tried to do this above, and have the
commented out original below}

@@

%%%%%%%%%%
\section{Applications} 
Hoekstra, Nachman, and colleagues have made an extensive study of coat color in the rock pocket mouse, \emph{Chaetodipus intermedius}, building on classic work by \citet{dice1940ecologic,benson1933concealing}. These mice live on rock outcrops throughout the southwestern United States and northern Mexico, 
and generally have light pigmentation, presumably to match the color of the rock, avoiding visual-based predators. 
However, in some regions these mice live on darker substrates such as old lava flows, 
and in these patches have much darker pigmentation. 
\citet{nachman2003different} demonstrated that on one of these flows (Pinacate), an allele at the MC1R locus is responsible for much of the change to a dark pelage.
This dark allele differs from the light allele by four amino acid changes, 
and has a dominant or partially dominant effect depending on the measure of coat color assessed. 
These lava flows are separated from each other by a range of distances, 
with some being separated from any other by hundreds of kilometers of light colored substrate. 
The Pinacate allele is not present in a number of other populations of dark colored rock pocket mouse, 
suggesting these populations have adapted in parallel. 
However, there is some evidence \citep{Hoekstra:05} that, elsewhere in the range, 
multiple dark outcrops may share a dark phenotype whose genetic basis has been spread by migration
despite intervening light habitat.

@@

The mutational target size $\mu$ for the trait is unclear. 
While the Pinacate dark haplotype differs from the light haplotype at three amino-acid residues,
it is likely that not all of these changes are needed for a population to begin to  
adapt. Also there a number of genes, beyond MC1R, at which adaptative changes affecting 
pigmentation have been identified in closely related species and more
broadly across vertebrates \citep{REF}.To span a range of plausible
values we use a low mutation rate of $\mu= 10^{-8}$, where the adaptive
mtation has to occur at a single base pair; and a high mutation
rate $\mu= 10^{-5}$, equivalent to the mutation being able to arise at any of a thousand base pairs
e.g. any of the coding bases of a typical length gene.  We set $A=100km^2$ reflecting a patch with a width of $\sim 10km$.

>>>>

  new mutation: \eqref{eqn:mutrate}; 
 via migration  \eqref{eqn:migrate}

Expected transit time \label{eqn:mean_tau}
Expected haplotype length \label{eqn:haplotype_length}

@@

Taking these approximations at face value, 
and assuming both rates are small, 
the time until the first of the two patches is colonized by the $B$
allele will be approximately exponentially distributed with mean $1/(2
\mutrate)$.
Following this, the time until the second patch is subsequently colonized 
(via either migration or new mutation) 
will be exponentially distributed with mean $1/(\mutrate+\migrate)$.
Both these rates scale linearly with population density ($\rho$) 
and the probability of establishment of a rare allele ($p_e\approx 2 s_b/\xi^2$),
so that the overall time to adaptation is robustly predicted to increase with offspring variance ($\xi^2$)
and decrease with population density and benefical selection coefficient.
Furthermore, the probability that the second adaptation is a new mutation,
i.e.\ the probability of parallel adapatation, is in $d=1$ ($d=2$ only differs slightly)
\begin{equation} \label{eqn:parallel_prob}
  \frac{\mutrate}{\mutrate+\migrate(r)} = \frac{ A \mu \alpha }{2 A
    \mu \alpha + \exp\left(- \sqrt{2 s_m} r / \sigma \right) },  
\end{equation}
so that the probability of parallel mutation should increase
approximately logistically with the distance between the patches, at rate $\sqrt{s_m} /\sigma$.

@@

If not yet adapted, a patch of area $A$ will acquire the adaptation by new mutation at rate $2 s_p \mu \rho A/\xi^2$ (equation \eqref{eqn:mutrate}).
Suppose that patches $1, \ldots, k$ are already adapted,
and that these patches are at distances $r_1, \ldots, r_k$ away from this unadapted patch.
The total rate of adaptation through migration, from equation \eqref{eqn:migrate}, is
\begin{equation}
  \migrate(r_i) = C \frac{ \rho s_p }{\xi^2 s_m} \sum_{i=1}^{k} \left(r_i \sqrt{2 s_m} /\sigma \right)^{-(d-1)/2} \exp\left(- r_i \sqrt{2 s_m} /\sigma\right),
\end{equation}
and if the patch adapts through migration, the probability the adapted allele 
comes from patch $i$ is 
\begin{align}  \label{eqn:prob-parallel}
  \frac{\left(r_i \right)^{-(d-1)/2} \exp\left(- r_i \sqrt{2 s_m}
  /\sigma\right)} {\sum_{j=1}^{k}  \left(r_i \right)^{-(d-1)/2} \exp\left(- r_i \sqrt{2 s_m}
    /\sigma\right) } .
\end{align}
These specify the transition rates of the process.

@@

@@

%%%%%%%%%%
\section{Applications} 
Hoekstra, Nachman, and colleagues have made an extensive study of coat color in the rock pocket mouse, \emph{Chaetodipus intermedius}, building on classic work by \citet{dice1940ecologic,benson1933concealing}. These mice live on rock outcrops throughout the southwestern United States and northern Mexico, 
and generally have light pigmentation, presumably to match the color of the rock, avoiding visual-based predators. 
However, in some regions these mice live on darker substrates such as old lava flows, 
and in these patches have much darker pigmentation. 
\citet{nachman2003different} demonstrated that on one of these flows (Pinacate), an allele at the MC1R locus is responsible for much of the change to a dark pelage.
This dark allele differs from the light allele by four amino acid changes, 
and has a dominant or partially dominant effect depending on the measure of coat color assessed. 
These lava flows are separated from each other by a range of distances, 
with some being separated from any other by hundreds of kilometers of light colored substrate. 
The Pinacate allele is not present in a number of other populations of dark colored rock pocket mouse, 
suggesting these populations have adapted in parallel. 
However, there is some evidence \citep{Hoekstra:05} that, elsewhere in the range, 
multiple dark outcrops may share a dark phenotype whose genetic basis has been spread by migration
despite intervening light habitat. We use our framework to investigate
how the probability of parallel mutation depends on the distance
between lava flow patches.

@@

The mutational target size $\mu$ for the trait is unclear. 
While the Pinacate dark haplotype differs from the light haplotype at three amino-acid residues,
it is likely that not all of these changes are needed for a population to begin to  
adapt. Also there a number of genes, beyond MC1R, at which adaptative changes affecting 
pigmentation have been identified in closely related species and more
broadly across vertebrates \citep{REF}.To span a range of plausible
values we use a low mutation rate of $\mu= 10^{-8}$, where the adaptive
mtation has to occur at a single base pair; and a high mutation
rate $\mu= 10^{-5}$, equivalent to the mutation being able to arise at any of a thousand base pairs
e.g. any of the coding bases of a typical length gene.  Finally we set
$A=100km^2$ reflecting a patch with a width of $\sim 10km$. In Figure
\ref{fig:mice_prob_parallel} we show the dependance of the probability
of parallel mutation on the distance between lava flow patches using
these parameters. As can be seen parallel mutation should become
likely over a scale of tens to a few hundred kilometers between patches.

\begin{figure}[ht]
  \begin{center}
    \includegraphics{Lava_flow_mice_prob_parallel}
  \end{center}
  \caption{
The probability of parallel mutation as a function of the distance
between environmental patches for two different cline widths
and two different mutation rates (using eqn. \eqref{eqn:prob-parallel}). The parameters were chosen to match  
  \label{fig:mice_prob_parallel}
  }
\end{figure}
